Choosing Color Palettes for Statistical Graphics

### 2006-09-07 trunk/vcd/paper/Color/draft.Rnw
recompiled color draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@546 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
improved captions
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@559 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

generate_png <- FALSE
@

@@

\includegraphics{hcl-dcone}
\caption{\label{fig:hsv-hcl} HSV and HCL space}
\end{center}
\end{figure}

@@

\includegraphics{hcl-dcone-gray}
\caption{\label{fig:hsv-hcl-gray} HSV and HCL space (in gray levels)}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-q-construct>>=
if(generate_png) {
  png(file = "pal-q-construct.png", height = 480, width = 480)
  hcl.wheel(luminance = 70, do.label = FALSE, inner.circle = TRUE)
  rect(0.45, -0.05, 0.55, 0.05, col = hcl(0, 50, 70))
  rect(-cos(pi/3)/2-0.05, sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, sin(pi/3)/2+0.05, col = hcl(120, 50, 70))
  rect(-cos(pi/3)/2-0.05, -sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, -sin(pi/3)/2+0.05, col = hcl(240, 50, 70))
  dev.off()
}
@
\includegraphics{pal-q-construct}
\vspace*{-2cm}
\caption{\label{fig:pal-q-construct} Constructing qualitative palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-q-examples, fig=TRUE, height=4, width=6>>=
par(mfrow = c(2, 2), mar = c(0, 0, 3, 0))
pal(rainbow_hcl(4, start = 30, end = 300), main = "dynamic [30, 300]")
pal(rainbow_hcl(4, start = 60, end = 240), main = "harmonic [60, 240]")
pal(rainbow_hcl(4, start = 270, end = 150), main = "cold [270, 150]")
pal(rainbow_hcl(4, start = 90, end = -30), main = "warm [90, -30]")
@
\caption{\label{fig:pal-q-examples} Examples for qualitative palettes}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<pal-s-construct>>=
if(generate_png) {
  png(file = "pal-s-construct.png", height = 420, width = 840)
  par(mfrow = c(1, 2))
  hue.slice(260, grid.n = 301)
  lines(c(0, 100), c(90, 50), lty = 2)
  lines(c(0, 80), c(90, 30), lty = 2)
  lines(c(2.5, 2.5), c(90, 30), lty = 2)
  rect(0, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(47.5, 67.5, 52.5, 72.5, col = hcl(260, 50, 70))
  rect(95, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  rect(37.5, 57.5, 42.5, 62.5, col = hcl(260, 40, 60))
  rect(77.5, 27.5, 82.5, 32.5, col = hcl(260, 80, 30))
  rect(0, 57.5, 5, 62.5, col = hcl(260, 0, 60))
  rect(0, 27.5, 5, 32.5, col = hcl(260, 0, 30))
  hue.slice(120, grid.n = 301)
  lines(c(0, 100), c(90, 75), lty = 2)
  lines(c(2.5, 2.5), c(90, 30), lty = 2)
  rect(0, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(47.5, 80, 52.5, 85, col = hcl(120, 50, 82.5))
  rect(95, 72.5, 100, 77.5, col = hcl(120, 100, 75))
  rect(0, 57.5, 5, 62.5, col = hcl(260, 0, 60))
  rect(0, 27.5, 5, 32.5, col = hcl(260, 0, 30))
  dev.off()
}
@
\includegraphics{pal-s-construct}
\caption{\label{fig:pal-s-construct} Constructing sequential palettes}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-s-examples, fig=TRUE, height=9, width=6>>=
par(mfrow = c(5, 1), mar = c(0, 0, 0, 0))
pal(sequential_hcl(12, c = 0, power = 2.2))
pal(sequential_hcl(12, power = 2.2))
pal(terrain_hcl(12, c = c(65, 0), l = c(45, 95), power = c(1/3, 1.5)))
pal(heat_hcl(12, c = c(80, 30), l = c(30, 90), power = c(1/5, 1.5)))
pal(rev(heat_hcl(12, h = c(0, -100), l = c(75, 40), c = c(40, 80), power = 1)))
@
\caption{\label{fig:pal-s-examples} Examples for sequential palettes}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<pal-d-construct>>=
if(generate_png) {
  png(file = "pal-d-construct.png", height = 480, width = 540)
  hue260 <- hue.slice(260, grid.n = 301, plot = FALSE)
  hue360 <- hue.slice(360, grid.n = 301, plot = FALSE)
  mychroma <- as.numeric(colnames(hue260))
  luminance <- as.numeric(rownames(hue260))
  nc <- ncol(hue260)
  nl <- nrow(hue260)
  plot(0.5, 0.5, xlim = c(-100, 100), ylim = c(0, 100), type = "n", axes = FALSE,
       xlab = "", ylab = "", yaxs = "i", xaxs = "i", main = "")
  for(i in 1:(nc-1)) {
    rect(mychroma[i], luminance[-nl], mychroma[i] + 100/(nc-1), luminance[-1], border = hue260[,i+1], col = hue260[,i+1])
  }
  for(i in 1:(nc-1)) {
    rect(-mychroma[i], luminance[-nl], -(mychroma[i] + 100/(nc-1)), luminance[-1], border = hue360[,i+1], col = hue360[,i+1])
  }
  axis(1, labels = FALSE, at = c(-100, 0, 100))
  axis(2, at = c(0, 30, 50, 60, 70, 90, 100), labels = c(0, 30, 50, 60, 70, 90, 100))
  axis(3, at = -4:4*25, labels=c(4:0*25, 1:4*25))
  axis(4, at = c(0, 30, 50, 60, 70, 90, 100), labels = c(0, 30, 50, 60, 70, 90, 100))
  mtext(c("hue = 0", "hue = 260"), side = 3, at = c(-50, 50), line = 3, cex = 1.2)
  mtext("chroma", side = 3, at = 0, line = 2)
  mtext("luminance", side = 2, at = 50, line = 2)
  mtext("luminance", side = 4, at = 50, line = 2)
  lines(c(-100, 0, 100), c(50, 90, 50), lty = 2)
  rect(-100, 47.5, -90, 52.5, col = hcl(0, 100, 50))
  rect(-55, 67.5, -45, 72.5, col = hcl(0, 50, 70))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(45, 67.5, 55, 72.5, col = hcl(260, 50, 70))
  rect(90, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  lines(c(-80, 0, 80), c(30, 90, 30), lty = 2)
  rect(-85, 27.5, -75, 32.5, col = hcl(0, 80, 30))
  rect(-45, 57.5, -35, 62.5, col = hcl(0, 40, 60))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(35, 57.5, 45, 62.5, col = hcl(260, 40, 60))
  rect(75, 27.5, 85, 32.5, col = hcl(260, 80, 30))
  dev.off()
}
@
\includegraphics{pal-d-construct}
\caption{\label{fig:pal-d-construct} Constructing diverging palettes}
\end{center}
\end{figure}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-d-examples, fig=TRUE, height=8, width=6>>=
par(mfrow = c(4, 1), mar = c(0, 0, 0, 0))
pal(diverge_hcl(7))
pal(diverge_hcl(7, c = 100, l = c(50, 90), power = 1))
pal(diverge_hcl(7, h = c(180, 330), c = 59, l = c(75, 95)))
pal(diverge_hcl(7, h = c(130, 43), c = 100, l = c(70, 90)))
@
\caption{\label{fig:pal-d-examples} Examples for diverging palettes}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<bkde, height=6, width=12>>=
if(generate_png) {
  png(file = "bkde.png", height = 420, width = 840)
  par(mfrow = c(1, 2))
  data("geyser", package = "MASS")
  dens <- bkde2D(cbind(geyser$duration, geyser$waiting), bandwidth = c(0.2, 3), gridsize = c(201, 201))
  image(dens$x1, dens$x2, sqrt(dens$fhat), xlab = "duration", ylab = "waiting", 
    col = rev(sequential_hcl(33, c = 0, power = 2)))
  box()
  image(dens$x1, dens$x2, sqrt(dens$fhat), xlab = "duration", ylab = "waiting", 
    col = rev(heat_hcl(33, c = c(80, 30), l = c(30, 90), power = c(1/5, 2))))
  box()
  dev.off()
}
@
\includegraphics{bkde}
\caption{\label{fig:bkde} Bivariate density estimation for Old Faithful geyser eruptions}
\end{center}
\end{figure}

\setkeys{Gin}{width=.6\textwidth}
\begin{figure}[p]
\begin{center}
<<seats, fig=TRUE, height=5, width=6>>=
par(mar = rep(0.8, 4))
pie(seats, clockwise = TRUE, col = hcl(c(240, 60, 300, 120, 0), 60, 75), radius = 1)
@
\caption{\label{fig:seats} Seats in the German parliament}
\end{center}
\end{figure}

@@

mosaic(Votes, labeling_args = list(rot_labels = c(0, 90, 0, 0), pos_labels = "center",
  just_labels = c("center", "center", "center", "right")),
  gp = gpar(fill = hcl(c(240, 60, 0, 120, 300), 60, 75)),
  spacing = spacing_highlighting, labeling = labeling_left,
  margins = unit(c(2.5, 1, 1, 12), "lines"), keep = FALSE)
@
\caption{\label{fig:votes} Votes in the German election 2005}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=.7\textwidth}
\begin{figure}[p]
\begin{center}
<<arthritis, fig=TRUE, height=6, width=7>>=
art <- xtabs(~ Treatment + Improved, data = Arthritis, subset = Sex == "Female")
names(dimnames(art))[2] <- "Improvement"
set.seed(1234)
mosaic(art, gp = shading_max)
@
\caption{\label{fig:arthritis} Extended mosaic display for arthritis data}
\end{center}
\end{figure}

\setkeys{Gin}{width=.7\textwidth}
\begin{figure}[p]
\begin{center}
<<class, height=6, width=8>>=
if(generate_png) {
  png(file = "class.png", height = 480, width = 640)
  set.seed(123)
  x1 <- rmvnorm(75, mean = c(1.5, 1.5), sigma = matrix(c(1, 0.8, 0.8, 1), ncol = 2))
  x2 <- rmvnorm(75, mean = c(-1, -1), sigma = matrix(c(1, -0.3, -0.3, 1), ncol = 2))
  X <- rbind(x1, x2)
  ex1 <- data.frame(class = factor(c(rep("a", 75), rep("b", 75))), x1 = X[,1], x2 = X[,2])
  rm(x1, x2, X)
  fm <- ksvm(class ~ ., data = ex1, C = 0.5)
  plot(fm, data = ex1)
  dev.off()
}
@
\includegraphics{class}
\caption{\label{fig:class} SVM classification plot}
\end{center}
\end{figure}

@@

>>>>

generate_png <- TRUE
@

@@

\includegraphics{hcl-dcone}
\caption{\label{fig:hsv-hcl} HSV and HCL space. The circles show the hue/saturation
and hue/chroma plane, respectively, for varying levels of value and luminance.}
\end{center}
\end{figure}

@@

\includegraphics{hcl-dcone-gray}
\caption{\label{fig:hsv-hcl-gray} HSV and HCL space (in gray levels). The circles
show the hue/saturation and hue/chroma plane, respectively, for varying levels of
value and luminance.}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-q-construct>>=
if(generate_png) {
  png(file = "pal-q-construct.png", height = 480, width = 480)
  hcl.wheel(luminance = 70, do.label = FALSE, inner.circle = TRUE)
  rect(0.45, -0.05, 0.55, 0.05, col = hcl(0, 50, 70))
  rect(-cos(pi/3)/2-0.05, sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, sin(pi/3)/2+0.05, col = hcl(120, 50, 70))
  rect(-cos(pi/3)/2-0.05, -sin(pi/3)/2-0.05, -cos(pi/3)/2+0.05, -sin(pi/3)/2+0.05, col = hcl(240, 50, 70))
  dev.off()
}
@
\includegraphics{pal-q-construct}
\vspace*{-2cm}
\caption{\label{fig:pal-q-construct} Constructing qualitative palettes. In the hue/chroma
plane for $L = 70$, the dashed circle correponds to a radius $C = 50$ with chosen angles
$H = 0, 120, 240$.}
\end{center}
\end{figure}

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-q-examples, fig=TRUE, height=4, width=6>>=
par(mfrow = c(2, 2), mar = c(0, 0, 3, 0))
pal(rainbow_hcl(4, start = 30, end = 300), main = "dynamic [30, 300]")
pal(rainbow_hcl(4, start = 60, end = 240), main = "harmonic [60, 240]")
pal(rainbow_hcl(4, start = 270, end = 150), main = "cold [270, 150]")
pal(rainbow_hcl(4, start = 90, end = -30), main = "warm [90, -30]")
@
\caption{\label{fig:pal-q-examples} Examples for qualitative palettes. Hue is varied
in different intervals for given $C = 50$ and $L = 70$.}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<pal-s-construct>>=
if(generate_png) {
  png(file = "pal-s-construct.png", height = 420, width = 840)
  par(mfrow = c(1, 2))
  hue.slice(260, grid.n = 301, main = "H = 260")
  lines(c(0, 100), c(90, 50), lty = 2)
  lines(c(0, 80), c(90, 30), lty = 2)
  lines(c(2.5, 2.5), c(90, 30), lty = 2)
  rect(0, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(47.5, 67.5, 52.5, 72.5, col = hcl(260, 50, 70))
  rect(95, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  rect(37.5, 57.5, 42.5, 62.5, col = hcl(260, 40, 60))
  rect(77.5, 27.5, 82.5, 32.5, col = hcl(260, 80, 30))
  rect(0, 57.5, 5, 62.5, col = hcl(260, 0, 60))
  rect(0, 27.5, 5, 32.5, col = hcl(260, 0, 30))
  hue.slice(120, grid.n = 301, main = "H = 120")
  lines(c(0, 100), c(90, 75), lty = 2)
  lines(c(2.5, 2.5), c(90, 30), lty = 2)
  rect(0, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(47.5, 80, 52.5, 85, col = hcl(120, 50, 82.5))
  rect(95, 72.5, 100, 77.5, col = hcl(120, 100, 75))
  rect(0, 57.5, 5, 62.5, col = hcl(260, 0, 60))
  rect(0, 27.5, 5, 32.5, col = hcl(260, 0, 30))
  dev.off()
}
@
\includegraphics{pal-s-construct}
\caption{\label{fig:pal-s-construct} Constructing sequential palettes. The chroma/luminance
plane is show for two hues $H = 260$ (left) and $H = 120$ (right). Colors are chosen by varying
either only luminance or both luminance and chroma.}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-s-examples, fig=TRUE, height=9, width=6>>=
par(mfrow = c(5, 1), mar = c(0, 0, 0, 0))
pal(sequential_hcl(12, c = 0, power = 2.2))
pal(sequential_hcl(12, power = 2.2))
pal(terrain_hcl(12, c = c(65, 0), l = c(45, 95), power = c(1/3, 1.5)))
pal(heat_hcl(12, c = c(80, 30), l = c(30, 90), power = c(1/5, 1.5)))
pal(rev(heat_hcl(12, h = c(0, -100), l = c(75, 40), c = c(40, 80), power = 1)))
@
\caption{\label{fig:pal-s-examples} Examples for sequential palettes.}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<pal-d-construct>>=
if(generate_png) {
  png(file = "pal-d-construct.png", height = 480, width = 540)
  hue260 <- hue.slice(260, grid.n = 301, plot = FALSE)
  hue360 <- hue.slice(360, grid.n = 301, plot = FALSE)
  mychroma <- as.numeric(colnames(hue260))
  luminance <- as.numeric(rownames(hue260))
  nc <- ncol(hue260)
  nl <- nrow(hue260)
  plot(0.5, 0.5, xlim = c(-100, 100), ylim = c(0, 100), type = "n", axes = FALSE,
       xlab = "", ylab = "", yaxs = "i", xaxs = "i", main = "")
  for(i in 1:(nc-1)) {
    rect(mychroma[i], luminance[-nl], mychroma[i] + 100/(nc-1), luminance[-1], border = hue260[,i+1], col = hue260[,i+1])
  }
  for(i in 1:(nc-1)) {
    rect(-mychroma[i], luminance[-nl], -(mychroma[i] + 100/(nc-1)), luminance[-1], border = hue360[,i+1], col = hue360[,i+1])
  }
  axis(1, labels = FALSE, at = c(-100, 0, 100))
  axis(2, at = c(0, 30, 50, 60, 70, 90, 100), labels = c(0, 30, 50, 60, 70, 90, 100))
  axis(3, at = -4:4*25, labels=c(4:0*25, 1:4*25))
  axis(4, at = c(0, 30, 50, 60, 70, 90, 100), labels = c(0, 30, 50, 60, 70, 90, 100))
  mtext(c("hue = 0", "hue = 260"), side = 3, at = c(-50, 50), line = 3, cex = 1.2)
  mtext("chroma", side = 3, at = 0, line = 2)
  mtext("luminance", side = 2, at = 50, line = 2)
  mtext("luminance", side = 4, at = 50, line = 2)
  lines(c(-100, 0, 100), c(50, 90, 50), lty = 2)
  rect(-100, 47.5, -90, 52.5, col = hcl(0, 100, 50))
  rect(-55, 67.5, -45, 72.5, col = hcl(0, 50, 70))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(45, 67.5, 55, 72.5, col = hcl(260, 50, 70))
  rect(90, 47.5, 100, 52.5, col = hcl(260, 100, 50))
  lines(c(-80, 0, 80), c(30, 90, 30), lty = 2)
  rect(-85, 27.5, -75, 32.5, col = hcl(0, 80, 30))
  rect(-45, 57.5, -35, 62.5, col = hcl(0, 40, 60))
  rect(-5, 87.5, 5, 92.5, col = hcl(260, 0, 90))
  rect(35, 57.5, 45, 62.5, col = hcl(260, 40, 60))
  rect(75, 27.5, 85, 32.5, col = hcl(260, 80, 30))
  dev.off()
}
@
\includegraphics{pal-d-construct}
\caption{\label{fig:pal-d-construct} Constructing diverging palettes. The chroma/luminance
plane is shown back to back hues $H = 0$ and $H = 260$. Colors are chosen by simultaneously 
varying luminance and chroma.}
\end{center}
\end{figure}

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<pal-d-examples, fig=TRUE, height=8, width=6>>=
par(mfrow = c(4, 1), mar = c(0, 0, 0, 0))
pal(diverge_hcl(7))
pal(diverge_hcl(7, c = 100, l = c(50, 90), power = 1))
pal(diverge_hcl(7, h = c(180, 330), c = 59, l = c(75, 95)))
pal(diverge_hcl(7, h = c(130, 43), c = 100, l = c(70, 90)))
@
\caption{\label{fig:pal-d-examples} Examples for diverging palettes.}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[b!]
\begin{center}
<<bkde, height=6, width=12>>=
if(generate_png) {
  png(file = "bkde.png", height = 420, width = 840)
  par(mfrow = c(1, 2))
  data("geyser", package = "MASS")
  dens <- bkde2D(cbind(geyser$duration, geyser$waiting), bandwidth = c(0.2, 3), gridsize = c(201, 201))
  image(dens$x1, dens$x2, sqrt(dens$fhat), xlab = "duration", ylab = "waiting", 
    col = rev(sequential_hcl(33, c = 0, power = 2)))
  box()
  image(dens$x1, dens$x2, sqrt(dens$fhat), xlab = "duration", ylab = "waiting", 
    col = rev(heat_hcl(33, c = c(80, 30), l = c(30, 90), power = c(1/5, 2))))
  box()
  dev.off()
}
@
\includegraphics{bkde}
\caption{\label{fig:bkde} Bivariate density estimation for Old Faithful geyser eruptions.}
\end{center}
\end{figure}

\setkeys{Gin}{width=.6\textwidth}
\begin{figure}[p]
\begin{center}
<<seats, fig=TRUE, height=5, width=6>>=
par(mar = rep(0.8, 4))
pie(seats, clockwise = TRUE, col = hcl(c(240, 60, 300, 120, 0), 60, 75), radius = 1)
@
\caption{\label{fig:seats} Seats in the German parliament.}
\end{center}
\end{figure}

@@

mosaic(Votes, labeling_args = list(rot_labels = c(0, 90, 0, 0), pos_labels = "center",
  just_labels = c("center", "center", "center", "right")),
  gp = gpar(fill = hcl(c(240, 60, 0, 120, 300), 60, 75)),
  spacing = spacing_highlighting, labeling = labeling_left,
  margins = unit(c(2.5, 1, 1, 12), "lines"), keep = FALSE)
@
\caption{\label{fig:votes} Votes in the German election 2005.}
\end{center}
\end{figure}

@@

\setkeys{Gin}{width=.7\textwidth}
\begin{figure}[p]
\begin{center}
<<arthritis, fig=TRUE, height=6, width=7>>=
art <- xtabs(~ Treatment + Improved, data = Arthritis, subset = Sex == "Female")
names(dimnames(art))[2] <- "Improvement"
set.seed(1234)
mosaic(art, gp = shading_max)
@
\caption{\label{fig:arthritis} Extended mosaic display for arthritis data.}
\end{center}
\end{figure}

\setkeys{Gin}{width=.7\textwidth}
\begin{figure}[p]
\begin{center}
<<class, height=6, width=8>>=
if(generate_png) {
  png(file = "class.png", height = 480, width = 640)
  set.seed(123)
  x1 <- rmvnorm(75, mean = c(1.5, 1.5), sigma = matrix(c(1, 0.8, 0.8, 1), ncol = 2))
  x2 <- rmvnorm(75, mean = c(-1, -1), sigma = matrix(c(1, -0.3, -0.3, 1), ncol = 2))
  X <- rbind(x1, x2)
  ex1 <- data.frame(class = factor(c(rep("a", 75), rep("b", 75))), x1 = X[,1], x2 = X[,2])
  rm(x1, x2, X)
  fm <- ksvm(class ~ ., data = ex1, C = 0.5)
  plot(fm, data = ex1)
  dev.off()
}
@
\includegraphics{class}
\caption{\label{fig:class} SVM classification plot.}
\end{center}
\end{figure}

@@

%FIXME% Acknowledgements
%% Paul Murrell, Ross Ihaka, Brian D. Ripley, David Meyer, Thomas Kneib, Michael Hoehle
